---
title: "rexplore"
output: html_notebook
---

```{r setup} 
library(DT)
library(tidyverse)
library(countrycode)
library(plotly)
#setwd("/home/project/src/data/raw")
setwd("/home/risto/gits/billionaires/src/data/raw")
data = read.csv('Billionaires_Statistics_Dataset.csv', na.strings=c("","NA"))
world.data = read.csv("world-data-2023.csv")
```

## Preprocessing
```{r}
#maybe just drop all of the billionaires without countries

#adding continent info to the dataset
continents = data.frame(country = unique(data$country))
data$continent = countrycode(sourcevar=data[, "country"], origin = "country.name", destination="continent")

```

## Analysis
### Billionaire Analysis
```{r analysis}
#billionaires by country
sort(table(data$country), decreasing = T)
```
- as expected, the us and china have the lions share of billionaires
```{r}
#by gender
prop.table(table(data$gender))
```
- also as expected, almost 90% of billionaires are men 

```{r}
#by industry
sort(table(data$industries), decreasing = T)
```
- nothing really interesting to note yet

```{r}
#by selfMade
prop.table(table(data$selfMade))
```
- an almost 70/30 split here in favor of being 'self-made'

```{r}
#looking at the intersecton of selfMade and industry
prop.table(table(data$selfMade, data$industries), margin=2)
```
- number of interesting things here:
- automotive, construction, fashion, food and beverage, and service are all pretty evenly split
- tech is a 93/7 split in favor of self-made, which makes a lot of sense
- telecom is also very heavily split in favor of self-made (84/16)
- all other ones are mostly a 70/30 split, with sports being a 64/36 split (which kina makes sense)


```{r}
#looking at billionaires by source
sort(table(data$source), decreasing = T)
```
- most common one is real estate and investments
- there is one 'used car' guy, which is really funny

```{r}
#looking at billionaires by country and industry
by.country.industries = data %>% select(country, industries) %>% count(country, industries)

#not sure there is enough billionaires per country to do much meaningful analysis on it 
by.country.industries
 
```

```{r}
#looking at billionaires by age
summary(data$age)
```
- mean and median are both 65

```{r}
#looking at billionaires by age and industry
#binning the ages by decade
by.age.industry = data %>% select(industries, age) %>% count(age, industries)


table(data$age, data$industries)

```

```{r}
#looking at billionaires by continent
table(data$continent)
```
- makes sense, proportionately more billionaires in americas and europe than in other continents

```{r}
#looking at number of billionaires by country population
table(data$population_country)
```

```{r}
#looking at billionaires and self made
boxplot(data$selfMade, log10(data$finalWorth))


```


```{r}
names(data)
```



### Country Analysis
```{r}
#looking at a histogram of all the tax rates
tax.rate = world.data$Total.tax.rate
tax.rate = as.numeric(gsub("%", "", tax.rate))
hist(tax.rate)

#looking at who has >100% tax rates
world.data[world.data$Total.tax.rate == "219.60%",]
world.data[world.data$Total.tax.rate == "106.30%",]

#removing those two tax rates and looking at the resulting histogram
hist(tax.rate[tax.rate < 100])

```
- comoros has a 219.60% tax rate, that seems incorrect. 
- argentina has a 106.30% tax rate, that also seems incorrect. 
- tax rate without the two large outliers is a pretty normal bell cuve 

```{r}
#looking at a pairs plot of the data 
pairs() # how to create a pairs plot

#TODO need to convert character values into numeric
world.pairs = world.data %>% select(Density..P.Km2., Agricultural.Land...., Land.Area.Km2., Armed.Forces.size, Birth.Rate, Co2.Emissions, CPI, CPI.Change...., Fertility.Rate, Forested.Area...., Gasoline.Price,  GDP, Gross.primary.education.enrollment...., Gross.tertiary.education.enrollment...., Infant.mortality, Life.expectancy, Maternal.mortality.ratio, Minimum.wage, Out.of.pocket.health.expenditure, Physicians.per.thousand, Population, Population..Labor.force.participation...., Tax.revenue...., Total.tax.rate, Unemployment.rate, Urban_population)


str(world.pairs)

```


```{r}
#TODO split out all these analyses into different code chunks
names(world.data)

```

## Basic modeling
```{r}
#creating a basic linear model to see if there are any countries that are over/under represented
#TODO need to create a column that contains how many billionaires are from each country 
#TODO will probably need to include all countries and set the ones that have no billionaires to 0



pop = data %>% group_by(country) %>% mutate(num.bill=n())

f = lm(num.bill~population_country, data=pop)

c = chisq.test(table(pop$num.bill, pop$country))
summary(c)
c$residuals
plot(c)

library(corrplot)
corrplot(c$stdres)
```

## Test
```{r}
#adding continents to the dataframe
test = data.frame(country = unique(data$country))
test$continent = countrycode(sourcevar=test[, "country"], origin = "country.name", destination="continent")


```

```{r}
#looking at the WDI package to add more info about other countries
library(WDI)

WDI(country="CA", indicator="NY.GDP.PCAP.KD", start=1960, end=NULL, extra=FALSE, cache=NULL, latest=NULL, language="en" )
```
